---
title: "Developing Data Products: R Markdown and Leaflet Project"
author: "Shayan (Sean) Taheri"
date: "June 10, 2019"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

```

<ul>
  <li> Each U.S. state flag is represented by each marker in the map. We place it at the approximating geographic center of each state. Two states of Alaska and Hawaii are positioned just off the West Coast for convenience (the *state.center* (a part of the *state* datasets from the *datasets* R package) is used for extraction of the data related to the longitude and the latitude of the center of each state). </li>
 
 <li> For observing the information of each state, it is required to click on a marker.
 
(i.e. the online sources such as [Wikipedia](https://en.wikipedia.org/wiki/List_of_U.S._states_and_territories_by_population) can be used for fetching the data regarding the population of each state. It is taken from the 2010 census). </li>
 
 <li> We can hover over the map to see the highlighted boundaries of each state and the state name. </li>
</ul>  
 

```{r, warning=FALSE, message=FALSE}

# Let's load the necessary packages:

library(leaflet)

library(maps)

library(RColorBrewer)

```


```{r}

# Generating the "state.db" data frame using the incorporated in the 
# "base" R package data sets - "state.abb", "state.center", "state.division",
# "state.region" and "state.x77":

state.db <- data.frame(state.abb, state.center[[1]], state.center[[2]], state.division, state.name, state.region, state.x77)

state.db$state.abb <- as.character(state.db$state.abb)
#class(state.db$state.abb)
state.db$state.name <- as.character(state.db$state.name)
#class(state.db$state.name)

ind <- match("state.region", colnames(state.db))

colnames(state.db)[1:ind] <- c("Abbreviation", "lng", "lat", "Division", "Name", "Region")

```


```{r}

# Generating the icon URLs for the 50 ".png" files with the flags of each 
# state (which we will later use as markers):

names <- gsub(pattern = " ", replacement = "_", x = state.name)


iconURL <- lapply(names, function(name){

     paste("C:/Users/shaya/Desktop/R_Markdown_and_Leaflet/flags/", name, ".png", sep = "")
})

listFlags <- icons(iconUrl = iconURL, iconWidth = 30, iconHeight = 30, iconAnchorX = 15, iconAnchorY = 15)

```


```{r}

# Using the package "maps" to generate a map of the states in the USA:

mapStates = map("state", fill = TRUE, plot = FALSE)

```


```{r}

# Creating the color palette for the map of the US states:

bins <- c(0, 500000, 4000000, 8000000, 15000000, 25000000, 30000000, 
          35000000, 40000000)

pal <- colorBin("YlOrRd", domain = state.db$Population, bins = bins)

```


```{r}

stateNames <- gsub(pattern = ":main", replacement = "", x = mapStates$names)

stateNames <- sapply(stateNames, function(name){
    
         paste(toupper(substr(x = name, 
                              start = 1, 
                              stop = 1)), 
                       substr(x = name, 
                              start = 2, 
                              stop = nchar(name)), 
                       sep = "")

    
})

stateNames <- as.vector(stateNames, mode = "character")


stateNamesUpper <- sapply(stateNames, function(name){
    
         if(gregexpr(pattern = " ", name)[[1]][1] > 0){
             
               paste(substr(x = name, 
                            start = 1,
                            stop = gregexpr(pattern = " ", name)[[1]][1]),
                     toupper(substr(x = name, 
                              start = gregexpr(pattern = " ", name)[[1]][1] + 1, 
                              stop = gregexpr(pattern = " ", name)[[1]][1] + 1)), 
                     substr(x = name, 
                            start = gregexpr(pattern = " ", name)[[1]][1] + 2, 
                            stop = nchar(name)), 
                     sep = "")

         }else paste(name)
    
})


stateNamesUpper <- as.vector(stateNamesUpper, mode = "character")

stateNamesUpper[8] <- "District Of Columbia" 



states <- c("California", "Texas", "Florida", "New York", "Pennsylvania",
            "Illinois", "Ohio", "Georgia", "North Carolina", "Michigan",
            "New Jersey", "Virginia", "Washington", "Arizona", "Massachusetts",
            "Tennessee", "Indiana", "Missouri", "Maryland", "Wisconsin",
            "Colorado", "Minnesota", "South Carolina", "Alabama", "Louisiana",
            "Kentucky", "Oregon", "Oklahoma", "Connecticut", "Iowa",
            "Utah", "Mississippi", "Arkansas", "Nevada", "Kansas",
            "New Mexico", "Nebraska", "West Virginia", "Idaho", "Hawaii",
            "New Hampshire", "Maine", "Rhode Island", "Montana", "Delaware", 
            "South Dakota", "North Dakota", "Alaska", "Vermont", "Wyoming")


population.current <- c(37254503, 25146105, 18804623, 19378087, 12702887,
                        12831549, 11536725, 9688681, 9535692, 9884129,
                        8791936, 8001045, 6724543, 6392307, 6547817, 6346275,
                        6484229, 5988927, 5773785, 5687289, 5029324, 
                        5303925, 4625401, 4780127, 4533479, 4339349,
                        3831073, 3751616, 3574118, 3046869, 2763888,
                        2968103, 2915958, 2700691, 2853132, 2059192,
                        1826341, 1853011, 1567652, 1360301, 1316466,
                        1328361, 1052931, 989417, 897936, 814191, 672591,
                        710249, 625745, 563767)


df <- data.frame(Name = states, Population.2010 = population.current)
df <- df[order(df$Name), ]
rownames(df) <- 1:nrow(df)
df$Name <- as.character(df$Name)


state.db <- merge(state.db, df, by = "Name")

mapStates$namesUpper <- stateNamesUpper
mapStates$namesUpper[23] <- "Michigan" 
mapStates$namesUpper[24] <- "Michigan" 


common.names <- intersect(df$Name, mapStates$namesUpper)
# Alaska, Hawaii
indices <- match(c("Alaska", "Hawaii"), df$Name)

```


```{r}

common.names <- intersect(df$Name, mapStates$namesUpper)

us.map <- data.frame(Name = mapStates$namesUpper)
us.map$Name <- as.character(us.map$Name)
us.map$Population <- vector(mode = "integer", length = nrow(us.map))

for(i in 1:nrow(us.map)){
 if(us.map$Name[i] %in% common.names){
    
    us.map$Population[i] <- df[df$Name == us.map$Name[i], ]$Population.2010
 }}

```


```{r}

info <- sapply(as.character(state.db$Name), function(name){
    
    paste(name, ",", "<br/>Abbreviation: ", state.db[state.db$Name == name, "Abbreviation"], "," , "<br/>Population: ", state.db[state.db$Name == name, "Population.2010"], ",", "<br/>Area: ", state.db[state.db$Name == name, "Area"], sep = "")
})

info <- as.vector(info, mode = "character")

```


```{r, warning=FALSE, message=FALSE}


state.db %>% 
  leaflet() %>%
  addTiles() %>%
  addMarkers(icon = listFlags, popup = info, options = markerOptions(opacity = 0.7)) %>%
addPolygons(lng = mapStates$x, lat = mapStates$y, 
            smoothFactor = 0.2,
            fillOpacity = 0.6, 
            fillColor = ~pal(us.map$Population), 
            stroke = TRUE, 
            dashArray = "3", 
            weight = 2, 
            color = "white", 
            opacity = 1,
            highlight = highlightOptions(weight = 5,
                                         color = "#666",
                                         dashArray = "",
                                         fillOpacity = 0.7,
                                         bringToFront = TRUE),
            label = us.map$Name,
            labelOptions = labelOptions(style = list("font-weight" = "normal", padding = "3px 8px"),
                           textsize = "15px",
                           direction = "auto")) %>%
addLegend(pal = pal, position = "topright", values = ~us.map$Population,
          title = "Population")


```